---
title: "Point Pattern Analysis"
author: "Cristina Martinez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   html_document:
    df_print: paged
    fig_height: 7
    fig_width: 7
#    toc: yes
#    toc_float: yes
always_allow_html: yes
#download_code: yes    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libs, message = FALSE, warning = FALSE}
options(Ncores = 12)
library(mapview)
library(sf)
library(censusxy)
library(dplyr)
library(tmap)
```

## Create point layer from data using latitude and longitude and plot points
```{r}
addr <- read.csv("C:/Users/chrys/Documents/GitHub/DEM7093/data/drinking_places_dt.csv")
addr <- addr %>%
  dplyr::select(names(addr)[c(4,6,14,23:24,41,42)])

results <- st_as_sf(addr, coords=c("Longitude", "Latitude"), crs=4269, agr="constant")

results.proj<-st_transform(results, crs = 2278)
```

```{r}
mv1 <- mapview(results.proj, layer.name = "Drinking Establishments")
#mapshot(mv1, file = paste0(getwd(), "/map1.png"))
mv1
```

```{r tmap1, eval = FALSE, include = FALSE}
library(tmap)
library(tmaptools)
library(OpenStreetMap)
bg <- read_osm(results.proj, ext=1.1)

tmap_mode("view")
tm_basemap("OpenStreetMap.Mapnik")+
  tm_shape(results.proj)+
  tm_dots(col = "blue")

tmap_mode("plot")
```

```{r tmap with markers, eval = FALSE, include = FALSE}
library(ggplot2)
library(ggmap)
library(ggspatial)
library(leaflet)
leaflet() %>%
  addTiles()%>%
  addMarkers(lng = addr$Longitude, lat= addr$Latitude)

```

# Mean feature - average of bar coordinates
Mean Center of the points is:
c(2121292.51040046, 13701736.8982003)
```{r}
mean_feature<-apply(st_coordinates(results.proj), MARGIN = 2, FUN = mean)
mean_feature<-data.frame(place="meanfeature", x=mean_feature[1], y= mean_feature[2])
mean_feature<-st_as_sf(mean_feature, coords = c("x", "y"), crs= 2278)


tmap_mode("view")
tm_basemap("OpenStreetMap.Mapnik")+
  tm_shape(results.proj, size = .2)+
  tm_dots(col = "purple")+
  tm_shape(mean_feature)+
  tm_dots(col = "green", size = .2)
```

```{r median of points, eval = FALSE, include = FALSE}
## Central feature - Median of coordinates
median_feature<-apply(st_coordinates(results.proj),
                      MARGIN = 2,
                      FUN = median)

median_feature<-data.frame(place="medianfeature",
                           x=median_feature[1],
                           y= median_feature[2])

median_feature<-st_as_sf(median_feature,
                         coords = c("x", "y"),
                         crs= 2278)


tmap_mode("view")
tm_basemap("OpenStreetMap.Mapnik")+
  tm_shape(results.proj)+
  tm_dots(col = "green", size = .2)+
  tm_shape(mean_feature)+
  tm_dots(col = "red", size=.2)+
  tm_shape(median_feature)+
  tm_dots(col = "blue", size = .2)

```
```{r buffer points, eval = FALSE, include = FALSE}
## Buffer points
bar_buff<- st_buffer(results.proj, dist = 1000)

#tmap_mode("plot")
tm_basemap("OpenStreetMap.Mapnik" )+
  tm_shape(results.proj, is.master = T)+
  tm_dots(col = "green")+
  tm_shape(bar_buff)+
  tm_polygons(col = "red", alpha = .1)

```

## Convex hull plot 
```{r convex hull 1}
chull <- st_convex_hull(st_union(results))

tmap_mode("view")
tm_basemap("OpenStreetMap.Mapnik")+
  tm_shape(results.proj)+
  tm_dots(col = "purple")+
  tm_shape(chull)+
  tm_polygons(col = "grey", alpha = .5)
```

## Alternative Convex Hull Plot
```{r convex hull 2}
mapview(chull, layer.name = "Convex Hull")+
  mapview(results, col.regions = "green", layer.name = "Drinking Establishments")
```

```{r kernel density, eval = FALSE, include = FALSE, message=FALSE, warning=FALSE}
## kernel density  - 
### You need projected data for this to work right

#R can do kernel density maps, but using simple features it's kind of complicated. Here is one way to do this using the `SpatialKDE` package:
library(SpatialKDE)

grid_groc <- results.proj %>%
  create_grid_rectangular(cell_size = 1000, side_offset = 2000)

kde <- results.proj%>%
  kde(band_width = 2000, kernel= "quartic", grid = grid_groc)

tm_shape(kde)+
  tm_polygons(col = "kde_value", palette= "viridis", title = "Density Estimate")+
  tm_shape(results.proj)+
  tm_dots()

```
## Spatial join 

```{r, results="hide"}
library(tidycensus)

#load census tract data
sa_acs<-get_acs(geography = "tract",
                state = "TX",
                county = "Bexar", 
                year = 2019,
                variables = "DP05_0001E",
                geometry = T, 
                output = "wide")

#rename variables and filter missing cases
sa_acs2<-sa_acs%>%
  mutate(totpop= DP05_0001E)%>%
  dplyr::select(GEOID, totpop)

sa_acs2<-st_transform(sa_acs2, crs = 2278)
sa_trol<-st_cast(sa_acs2, "MULTILINESTRING") #make outline
```

```{r}
spjoin<-st_join(results.proj, sa_acs2)
#head(spjoin)

tmap_mode("view")
tm_basemap("OpenStreetMap.Mapnik")+
  tm_shape(spjoin, is.master = T)+
  tm_dots("totpop", size = .1, title = "Population/Bar")+
  tm_shape(sa_acs2)+
  tm_polygons(alpha = .1)
```

## Count points in polygons 

```{r}
sa_acs2$nbar<- lengths(st_intersects(sa_acs2, results.proj))

sa_acs2$bar_pc <- 1000*(sa_acs2$nbar/sa_acs2$totpop)

tmap_mode("plot")
m<-tm_basemap("OpenStreetMap.Mapnik")+
  tm_shape(sa_acs2)+
  tm_polygons("bar_pc", title = "Bars / 1,000 People")+
  tm_shape(spjoin, is.master = T)+
  tm_dots( size = .07)

m
```



</br>



## Alternative Density Map 
```{r}
mapview(sa_acs2, zcol="bar_pc", layer.name = "Bars/1000 Pop")+
  mapview(results.proj, col.regions = "green", layer.name = "Bars")
```

## Nearest Neighbor analysis

Nearest neighbor analysis to see the distributions of space between features, a measure of clustering or dispersion.

```{r, message = FALSE, warning = FALSE, results = 'hide'}
library(spatstat)
bar.pp<-as.ppp(as(results.proj, "Spatial"))

plot(nearest.neighbour(bar.pp))
```

```{r, include = FALSE}
options(qgisprocess.path = "C://Program Files//QGIS 3.16.16//bin//qgis_process-qgis-ltr.bat" )
```

Average Nearest Neighbor (ANN) 1815/1885 = 0.963
<br/>ANN less than 1 indicates clustering 
</br>A z-score of -0.352 also indicates clustering and is not significant.

```{r qgis, warning = FALSE, message = FALSE, results = 'hide'}

library(qgisprocess)
qgis_configure()

bars_nn<-qgis_run_algorithm(alg="native:nearestneighbouranalysis",
                            INPUT=results.proj,
                            OUTPUT_HTML_FILE=file.path(tempdir(), "barsnn.html"),
                            load_output = TRUE)
```
```{r}
bars_nn
```


