---
title: "DEM 7093 Homework 3 using R"
author: "Cristina Martinez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   html_document:
   toc: yes
   toc_float: yes
always_allow_html: yes
---

```{r, message = F, warning = F}
library(tidycensus, quietly = T)
library(sf, quietly = T)
library(dplyr, quietly = T)
```

### Read in Bexar county tracts

```{r, results = "hide", message = F, warning = F}
sa_acs<-get_acs(geography = "tract",
                state = "TX",
                county = "Bexar",
                year = 2017,
                variables = c("DP05_0001E", "DP03_0119PE"),
                geometry = T, 
                output = "wide")

#create a county FIPS code - 5 digit
sa_acs$county<-substr(sa_acs$GEOID, 1, 5)

#rename variables and filter missing cases
sa_acs2<-sa_acs %>%
  mutate(totpop = DP05_0001E, ppov=DP03_0119PE) %>%
  na.omit()

# mydat<-sf::st_read("~/OneDrive - University of Texas at San Antonio/classes/gis_classwork/bexarpolygon.shp")
# plot(mydat)
```

### find coordinate system of current map

This shows that the current coordinate reference system is NAD83

```{r, warning = F}
st_crs(sa_acs2)
```

```{r, evaluate = F, include = F}
### create basic map
# library(tmap)
# library(tmaptools)
# 
# tm_shape(sa_acs2)+
#   tm_polygons("ppov", title="% in Poverty", palette="Blues", style="quantile", n=5 )+
#   tm_format("World", title="San Antonio Poverty Estimates - Quantile Breaks", legend.outside=T)+
#   tm_scale_bar()+
#   tm_compass()
```

### Un projected distance calculation

The distance between the centroids of the two tracts is 20,787.28 meters. This distance is not interpretable because it is not mapped in its actual location on the earth's surface.

```{r, warning = F}
new_sa<-st_transform(sa_acs2, crs = 2278)

#Extract two tracts
twtr<-new_sa%>%
  filter(GEOID %in% c(48029110100, 48029181820))

twtr<-sa_acs2%>%
  filter(GEOID %in% c(48029110100, 48029181820))

# get centroid coordinates for two tracts
tr_co<-st_centroid(twtr)

#Measure feet apart
st_distance(tr_co)
```

### Projected distance calculation

The distance between the centroids of the two tracts in its correct projection is 68,091.77 ft. We see the unit of measurement has changed. The actual difference in the measurement of the unprojected and projected measurements is about 100 ft, but running spatial analyses with the incorrect map projections can yield false results. Something may appear to be closer or farther than it actually is when using unprojected spatial data.

```{r, warning = F}
new_sa<-st_transform(sa_acs2, crs = 2278)

#Extract two tracts
twtr<-new_sa%>%
  filter(GEOID %in% c(48029110100, 48029181820))

# get centroid coordinates for two tracts
tr_co<-st_centroid(twtr)

#Measure feet apart
st_distance(tr_co)
```
```{r, evaluate = F, include = F}
# # Using QGIS within R
# This is another way to do the above task, by running a QGIS algorithm within R using the `qgisprocess` package. 
# 
# See the vignette [here](https://github.com/paleolimbot/qgisprocess) for more on what this package is and some examples.

# library(qgisprocess) #load the package
# 
# qgis_configure() #set up qgis - find the executable
# 
# # qgis_algorithms() lists all the available routines in QGIS
# head(qgis_algorithms())

```

```{r, evaluate = F, include = F}
#We can use `grep` to search for specific terms in the algorithms. 
# algs<-qgis_algorithms()
# 
# algs[grep(x = algs$algorithm, "distance"),"algorithm"]
# 
# qgis_show_help("qgis:distancematrix")
```

```{r, evaluate = F, include = F}
# out = qgis_run_algorithm(alg = "qgis:distancematrix",
#                INPUT = tr_co[1,],
#                INPUT_FIELD = "GEOID", 
#                TARGET = tr_co[2,],
#                TARGET_FIELD = "GEOID",
#                MATRIX_TYPE = 0, 
#                NEAREST_POINTS = 1)
# 
# output_sf <- sf::read_sf(qgis_output(out, "OUTPUT"))
# output_sf$Distance
```