---
title: "DEM 5093 7093 Lab 1 using  R - Basic Map making"
author: "Cristina Martinez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
always_allow_html: yes
---


```{r, message=FALSE, eval=TRUE, include = FALSE}
library(tidycensus)
library(tidyverse)
library(sf)
library(ggplot2)
library(classInt)
```

### Search for variables by keywords in the label 
```{r, eval = FALSE, echo = FALSE, include = FALSE}
v15_Profile%>%
  filter(grepl(pattern = "POVERTY", x = label))%>%
  select(name, label)
```

\newpage
## Extract from ACS summary file data

```{r, results='hide'}
sa_acs<-get_acs(geography = "tract",
                state= "TX",
                county = "Bexar",
                year = 2019,
                variables=c(medinc = "DP03_0062E", medinc.moe = "DP03_0062M"),
                geometry = T,
                output = "wide")

```

Here, I create some other variables that we may need later
```{r}
# create a county FIPS code - 5 digit and filter missing cases
sa_acs2 <- sa_acs %>%
  mutate(county = substr(GEOID, 1, 5)) %>%
  st_transform(crs = 2919)%>%
  na.omit()

```

## Write data out to shapefile
You may need to create or modify some data in R and then use it in the desktop GIS (QGIS), we can write any data from R into a variety of data formats using the  `sf::st_write()` function.


```{r, message=FALSE, warning=FALSE, eval=FALSE, include=FALSE}
#change the directory for your computer
sf::st_write(sa_acs2,
             dsn="../data/sa_tr_dp03",
             layer="sa_tr_dp03",
             driver="GPKG") ## Save as geopackage format - QGIS likes this

```

\newpage

## Some basic mapping of variables
Here I generate a quantile break for % black in census tracts and compare it to a Jenks break. *Note* in ggplot, the Jenks break is harder to do
```{r, message=FALSE, fig.width=12, fig.height=8}

library(patchwork)
library(ggsn)
library(XploreR)
source("https://raw.githubusercontent.com/coreysparks/Rcode/master/mutate_map_funs.R")

medinc_map<-sa_acs2 %>%
  mutate_map_brks(medinc, n=6, style="quantile") %>%
  mutate_map_brks(medinc, n=6, style="jenks")



p1<-ggplot(medinc_map, aes(fill = medinc_qbrks)) + 
  geom_sf(color="grey50") + 
  ggtitle("Median Household Income", 
          subtitle = "Bexar County Texas, 2019 - Quantile Breaks")+
    scale_fill_brewer(palette = "Blues") + 
  scale_color_brewer(palette = "Blues")+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank())+
  north(medinc_map)+
  scalebar(medinc_map, location="bottomleft",
           dist=5, transform = T,
           dist_unit = "km",
           model="WGS84",
           st.size =2 )
p1



p2<-ggplot(medinc_map, aes(fill = medinc_jbrks)) + 
  geom_sf(color="grey50") + 
  ggtitle("Median Household Income", 
          subtitle = "Bexar County Texas, 2019 - Jenks Breaks")+
  scale_fill_brewer(palette = "Blues") + 
  scale_color_brewer(palette = "Blues")+
    theme(axis.text.x = element_blank(),
          axis.text.y = element_blank())+
  north(medinc_map)+
  scalebar(medinc_map, 
           location="bottomleft",
           dist=5,
           transform = T,
           dist_unit = "km",
           model="WGS84",
           st.size =2)
p2
  
p1+ p2

```

You can save the image from above to your computer by using `ggsave()`
```{r, eval=FALSE}
ggsave(filename="../images/lab1map1.png",
       dpi = "print")

```

### Slightly easier way using `tmap`

The `tmap` package is an attractive alternative to using `ggplot()` when making maps, and makes basic cartographic principles easier.

*Note* `style="fisher"` is equivalent to `style="jenks"` and scales better to larger data.

```{r}
#library(tmap)
#library(tmaptools)

tm_shape(sa_acs2)+
  tm_polygons("medinc",
              title="Income",
              palette="Blues",
              style="quantile", 
              n=5, 
              legend.hist=T)+
  tm_format("World",
            title="San Antonio Median Household Income - Quantile Breaks",
            legend.outside=T)+
  tm_scale_bar()+
  tm_compass()

tm_shape(sa_acs2)+
  tm_polygons("medinc",
              title="Income",
              palette="Blues",
              style="fisher",
              n=5,
              legend.hist=T )+
  tm_format("World",
            title="San Antonio Median Household Income - Jenks Breaks",
            legend.outside=T)+
  tm_scale_bar()+
  tm_compass()
```


## Interactive map with mapview

```{r}
library(mapview)
library(RColorBrewer)

medinc_map$medinc_jbrks<-relevel(medinc_map$medinc_jbrks,ref = "0-10463" )

medinc_map2 <- medinc_map %>%
  select(Name = NAME, `Median Income` = medinc, MOE = medinc.moe, `Jenk Breaks` = medinc_jbrks)

#pal <- colorRampPalette(brewer.pal(7, "Blues")) #set colors

x<-mapview(medinc_map2,
        zcol="Jenk Breaks",
        legend=T,
        map.types="OpenStreetMap",
        layer.name="Income")

x

mapshot(x, file = "C:/Users/chrys/Documents/UTSA/7093 GIS/map.jpeg", remove_controls = c("zoomControl", "layersControl"))

```



